### 2017-07-19 preprocessing.Rmd

---
title: "Preprocessing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Quality Control

```{r}
library(bigsnpr)

NCORES <- parallel::detectCores() / 2
plink <- "../plink_linux_x86_64/plink"

popresQC.bed <- snp_plinkQC(
  plink.path = plink,
  prefix.in = "../POPRES_data/POPRES_allchr",
  file.type = "--file", # ped/map
  geno = 0.05,
  mind = 0.05,
  maf = 0.05,
  hwe = 1e-10,
  autosome.only = TRUE
)

popresQC2.bed <- snp_plinkIBDQC(plink, popresQC.bed, ncores = NCORES)
```

```{r}
celiacQC.bed <- snp_plinkQC(
  plink.path = plink,
  prefix.in = "../thesis-celiac/Dubois2010_data/FinnuncorrNLITUK1UK3hap300",
  geno = 0.05,
  mind = 0.05,
  maf = 0.05,
  hwe = 1e-10,
  autosome.only = TRUE
)

celiacQC2.bed <- snp_plinkIBDQC(plink, celiacQC.bed, ncores = NCORES)
```

## Transform PLINK files in big.matrix format

```{r}
popresQC.rds <- snp_readBed(popresQC2.bed, "popresQC")

system.time(
  celiacQC.rds <- snp_readBed(celiacQC2.bed, "celiacQC")
)
```

## Imputation of Celiac dataset

```{r}
system.time(
  celiac <- snp_attach(celiacQC.rds)
)
str(celiac, max.level = 2)
```

```{r}
G <- celiac$genotypes
big_counts(G, ind.col = 1:10)
```

```{r}

```