### 2017-07-26 GWAS-biobank-scale.Rmd
biobank scale

---
title: "Imputation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", 
                      fig.width = 15, fig.asp = 0.7)
```

## Form a larger dataset

```{r}
library(bigsnpr)

celiac <- snp_attach("backingfiles/celiacQC.rds") 
G <- celiac$genotypes

system.time(
  bedfile <- snp_writeBed(celiac, bedfile = "backingfiles/tmp_biobank.bed", 
                          ind.row = sort(rep_len(rows_along(G), 500e3)),
                          ind.col = sort(rep_len(cols_along(G), 1e6)))
)

# Size if the ".bed" file in GB
file.size(bedfile) / 1024^3
```

## Make a GWAS on this dataset

### Read the dataset in bigSNP format

```{r}
system.time(
  rdsfile <- snp_readBed(bedfile, backingfile = "tmp_biobank")
) # 1.6h
```

```{r}
bkfile <- sub("\\.rds$", ".bk", rdsfile)
# Size if the ".bk" file in GB
file.size(bkfile) / 1024^3
```

```{r}
system.time(
  biobank <- snp_attach(rdsfile)
)
```

### Clumping with MAF before PCA

```{r}
NCORES <- parallel::detectCores() / 2
# Get SNPs in long-range LD regions
system.time(
  ind.excl <- snp_indLRLDR(biobank$map$chromosome,
                           biobank$map$physical.pos)
)
# No parallelization for this huge dataset
system.time(
  ind.keep <- snp_clumping(biobank$genotypes, 
                           infos.chr = biobank$map$chromosome,
                           exclude = ind.excl,
                           ncores = 1)
) # 6.3h
```

### Computation of first 10 PCs

```{r}
# Parallelization is OK for PCA because 
# only a subset of this huge dataset is used 
system.time(
  obj.svd <- big_randomSVD(biobank$genotypes, 
                           fun.scaling = snp_scaleBinom(),
                           ind.col = ind.keep,
                           ncores = NCORES)
) # 2.4h
```

### GWAS on a continuous variable

```{r}
# Simulate some continous phenotype
y <- rnorm(nrow(biobank$genotypes))

# No parallelization for this huge dataset
system.time(
  gwas <- big_univLinReg(biobank$genotypes,
                         y.train = y,
                         covar.train = obj.svd$u, 
                         ncores = 1)
) # 3h
```

```{r}
snp_manhattan(gwas, biobank$map$chromosome, biobank$map$physical.pos,
              npoints = 50e3, coeff = 2)
```