### 2017-07-30 autoSVD.Rmd
autoSVD OK

---
title: "Auto SVD"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", 
                      fig.width = 16, fig.asp = 0.7, error = TRUE)
options(width = 85)
```

## Get the data

```{r}
suppressMessages({
  library(bigsnpr)
  library(tidyverse)
  library(foreach)
})

celiac <- snp_attach("backingfiles/celiacQC.rds")
G <- celiac$genotypes
CHR <- celiac$map$chromosome
POS <- celiac$map$physical.pos
NCORES <- parallel::detectCores() / 2
```

## Without removing long-range LD regions

```{r}
system.time(
  ind.keep <- snp_clumping(G, CHR, ncores = NCORES)
)
system.time(
  obj.svd <- big_randomSVD(G, snp_scaleBinom(), 
                           ind.col = ind.keep,
                           ncores = NCORES)
)
```

```{r, fig.asp=1}
(plot_loadingsA <- plot(obj.svd, type = "loadings", loadings = 1:6, coeff = 1))
```

```{r}
oneSNP <- which.max(abs(obj.svd$v[, 4]))
(plot_scoresA <- plot(obj.svd, type = "scores", scores = 4:5, coeff = 2) +
    aes(color = as.factor(attach.BM(G)[, ind.keep[oneSNP]])) + 
    labs(title = NULL, color = "Genotype") +
    theme(legend.position = c(0.15, 0.15)) +
    guides(color = guide_legend(override.aes = list(size = 4))))
```

## Automatically removing long-range LD regions

```{r}
system.time(
  test <- snp_autoSVD(G, CHR, POS, ncores = NCORES)
)
htmlTable::htmlTable(attr(test, "lrldr"), rnames = FALSE,
                     css.table = "margin-top: 1em; 
                                  margin-bottom: 1em;   
                                  margin-left: auto;
                                  margin-right: auto",
                     css.cell = "padding: 5px 5px 5px 5px")
```

```{r, comment=""}
attr(test, "lrldr") %>%
  transmute(
    Chromosome = Chr,
    `Start (Mb)` = Start / 1e6,
    `Stop (Mb)` = Stop / 1e6
  ) %>%
  xtable::xtable(digits = 1, caption = "", label = "", auto = TRUE)
```

```{r, fig.asp=1}
(plot_loadingsB <- plot(test, type = "loadings", loadings = 1:6, coeff = 1))
```

```{r}
# Get population from external files
pop.files <- list.files(path = "../thesis-celiac/Dubois2010_data/",
                        pattern = "cluster_*",
                        full.names = TRUE)
pop <- snp_getSampleInfos(celiac, pop.files)[[1]]
pop.names <- c("Netherlands", "Italy", "UK1", "UK2", "Finland")

(plot_scoresB <- plot(test, type = "scores", scores = 4:5, coeff = 2) +
    aes(color = pop.names[pop]) + 
    labs(title = NULL, color = "Population") +
    theme(legend.position = c(0.2, 0.75)) +
    guides(color = guide_legend(override.aes = list(size = 4))))
```

```{r}
# save plots for paper
p <- cowplot::plot_grid(plot_loadingsA, plot_loadingsB, align = "hv", 
                        ncol = 2, labels = c("A", "B"), label_size = 30)
ggsave("figures/loadings.png", width = 1620, height = 920, scale = 1/75)

p <- cowplot::plot_grid(plot_scoresA, plot_scoresB, align = "hv", 
                        ncol = 2, labels = c("A", "B"), label_size = 30)
ggsave("figures/scores.png", width = 1640, height = 750, scale = 1/75)
```

```{r, fig.asp=0.7, fig.width=18}
doParallel::registerDoParallel(NCORES)

nBoot <- 500

tribble(
  ~name,                                               ~svd,
  "When automatically removing long-range LD regions", test, 
  "When keeping long-range LD regions",                obj.svd
) %>%
  apply(1, function(method) {
    V <- method[["svd"]][["v"]]^2
    m <- nrow(V)
    foreach(ic = seq_len(nBoot), .combine = 'rbind') %dopar% {
      ind <- sort(sample(m, replace = TRUE))
      apply(V[ind, ], 2, ineq::Gini)
    } %>%
      reshape2::melt() %>%
      cbind(method = method[["name"]])
  }) %>%
  Reduce(f = 'rbind') %>%
  ggplot(aes(as.factor(Var2), value)) %>%
  bigstatsr:::MY_THEME(coeff = 2) +
  geom_hline(yintercept = 2 / pi, lty = 2) +
  geom_boxplot(aes(color = method), lwd = 1.5) +
  labs(y = "Gini coefficient of squared loadings", 
       x = "Principal Component", color = "") +
  theme(legend.position = "top") +
  guides(color = guide_legend(ncol = 1))

doParallel::stopImplicitCluster()

ggsave("figures/gini.png", width = 1341, height = 865, scale = 1/75)
```

## Removing long-range LD regions based on predefined table

```{r}
ind.excl0 <- snp_indLRLDR(CHR, POS)
ind.keep0 <- snp_clumping(G, CHR, exclude = ind.excl0, ncores = NCORES)
obj.svd0 <- big_randomSVD(G, snp_scaleBinom(), 
                          ind.col = ind.keep0,
                          ncores = NCORES)
```

```{r}
corr <- cor(test$u, obj.svd0$u)
rownames(corr) <- colnames(corr) <- paste0("PC", 1:10)
htmlTable::htmlTable(round(100 * corr, 2), 
                     align = paste(c("|", rep("c", 10)), collapse = ""),
                     css.table = "margin-top: 1em; 
                                  margin-bottom: 1em;   
                                  margin-left: auto;
                                  margin-right: auto",
                     css.cell = "padding: 5px 5px 5px 5px")
```

```{r}
mean(diag(abs(corr)))
```

```{r, include=FALSE}
library(gmailr)
send_message(
  mime(To = "florian.prive.21@gmail.com",
       From = "florian.prive.21@gmail.com",
       Subject = sprintf("Knitting of %s has finished",
                         knitr::current_input()))
)
```