### 2017-09-06 benchmark-pca.Rmd

---
title: "Benchmarking Principal Component Analysis software"
output: html_document
---

In this notebook, we will benchmark three different methods for computing first Principal Components (PCs) of a large genotype matrix.

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", 
                      fig.width = 16, fig.asp = 0.7, error = TRUE)
options(width = 85)
```

## Get the data

```{r}
suppressMessages({
  # devtools::install_github("privefl/bigsnpr")
  library(bigsnpr)   
  # devtools::install_github("gabraham/flashpca", subdir = "flashpcaR")
  library(flashpcaR)
  # install.packages("tidyverse")
  library(tidyverse)
  # install.packages("glue")
  library(glue)   
})
# FastPCA: https://data.broadinstitute.org/alkesgroup/EIGENSOFT/EIG-6.1.4.tar.gz
smartpca <- "../EIG-6.1.4/bin/smartpca"

NCORES <- nb_cores()

celiac <- snp_attach("backingfiles/celiacQC.rds")
G <- celiac$genotypes

# Thinning of SNPs
ind.excl <- snp_indLRLDR(celiac$map$chromosome,
                         celiac$map$physical.pos)
length(
  ind.keep <- snp_clumping(celiac$genotypes, celiac$map$chromosome,
                           exclude = ind.excl, 
                           ncores = NCORES)
)
```

```{r, include=FALSE}
# Some cleaning
file.remove(list.files(path = "backingfiles/", 
                       pattern = "celiacQC_keep*", 
                       full.names = TRUE))
```

```{r}
# Write new bed files keeping only a subset of SNPs and replicating individuals
snp_writeBed(celiac, "backingfiles/celiacQC_keep1.bed",
             ind.col = ind.keep)
snp_readBed("backingfiles/celiacQC_keep1.bed", "backingfiles/celiacQC_keep1")

snp_writeBed(celiac, "backingfiles/celiacQC_keep5.bed",
             ind.row = rep(rows_along(G), each = 5),
             ind.col = ind.keep)
snp_readBed("backingfiles/celiacQC_keep5.bed", "backingfiles/celiacQC_keep5")

snp_writeBed(celiac, "backingfiles/celiacQC_keep10.bed",
             ind.row = rep(rows_along(G), each = 10),
             ind.col = ind.keep)
snp_readBed("backingfiles/celiacQC_keep10.bed", "backingfiles/celiacQC_keep10")
```

## Functions for computing PCs

```{r}
# FlashPCA2: devtools::install_github("gabraham/flashpca", subdir = "flashpcaR")
flashpca <- function(prefix) {
  time <- system.time(
    svd <- flashpcaR::flashpca(prefix)
  )
  list(time[3], svd$vectors)
}

fastpca <- function(prefix, fast = TRUE) {
  parfile <- glue("{prefix}.pca.par")

  ## Renaming so that Eigensoft can recognize file type
  system(glue("cp {prefix}.bim {prefix}.pedsnp"))  
  ## Renaming so that Eigensoft can recognize file type
  system(glue("cp {prefix}.fam {prefix}.pedind"))  
  writeLines(glue(
    "genotypename: {prefix}.bed
    snpname: {prefix}.pedsnp
    indivname: {prefix}.pedind
    evecoutname: {prefix}.pca.evec
    evaloutname: {prefix}.eval
    altnormstyle: NO
    numoutevec: 10
    numoutlieriter: 0
    numoutlierevec: 10
    outliersigmathresh: 6
    qtmode: 0
    fastmode: {as.integer(fast)}"
  ), con = parfile)
  ## PCA
  time <- system.time(
    system(glue("{smartpca} -p {parfile}"), ignore.stdout = TRUE)
  ) 
  ## removing temp files
  system(glue("rm {prefix}.pedsnp")) 
  system(glue("rm {prefix}.pedind"))
  system(glue("rm {parfile}"))

  evec <- data.table::fread(glue("{prefix}.pca.evec"), 
                            data.table = FALSE, skip = 1)

  list(time[3], as.matrix(evec[, 2:11]))
}

# bigstatsr::big_randomSVD
mypca <- function(prefix, ncores = 1) {
  rdsfile <- paste0(prefix, ".rds")
  bigsnp <- snp_attach(rdsfile)

  time <- system.time(
    svd <- big_randomSVD(bigsnp$genotypes, 
                         fun.scaling = snp_scaleBinom(),
                         ncores = ncores)
  )

  list(time[3], svd$u)
}
```

## Evaluating all models

```{r}
# Preparing a dataset for the results
models <- tribble(
  ~method,               ~fun,     ~params,
  "FastPCA/EIGENSOFT",   fastpca,  list(),
  "FlashPCA2",           flashpca, list(),
  "bigstatsr - 1 core",  mypca,    list(),
  "bigstatsr - 6 cores", mypca,    list(ncores = NCORES)
) %>%
  mutate(size = list(c(1, 5, 10))) %>%
  unnest(size, .drop = FALSE) %>%
  mutate(time = NA, pcs = list(NA))

# Filling this data frame with time and PC scores for each method and dataset
for (i in rows_along(models)) { 
  size <- models$size[[i]]
  prefix <- glue("backingfiles/celiacQC_keep{size}")
  fun <- models$fun[[i]]
  params <- c(prefix, models$params[[i]])
  res <- do.call(fun, args = params)
  models[["time"]][[i]] <- res[[1]]
  models[["pcs"]][[i]] <- res[[2]]
}
```

## Plotting benchmark

```{r}
(models <- mutate(models, size = size * nrow(G)))
```

```{r}
models %>%
  ggplot(aes(size / 1000, time / 60, color = method)) %>%
  bigstatsr:::MY_THEME(coeff = 2.5) + 
  geom_point(aes(shape = method), cex = 8) +
  geom_line(aes(linetype = method), lwd = 3) + 
  theme(legend.position = c(0.25, 0.75),
        legend.key.width = unit(8, "line")) + 
  labs(x = "Number of individuals (x1000)", y = "Time (in minutes)", 
       color = "Methods:", linetype = "Methods:", shape = "Methods:")

ggsave("figures/benchmark-pca.png", width = 1271, height = 945, scale = 1/75)
```

## Comparing precision of results

```{r}
# SmartPCA
smartpca <- fastpca("backingfiles/celiacQC_keep1", fast = FALSE)
true <- smartpca[[2]]
cat(sprintf("Standard PCA takes forever: %.2gh.", smartpca[[1]] / 3600))
```

```{r}
# Tidy and plot results
(preci <- models %>% 
  mutate(
    size = as.factor(size),
    cor = lapply(pcs, function(x) {
      N <- nrow(x) / nrow(true)
      if (N == 1) {
        diag(cor(x, true))
      } else {
        diag(cor(x[c(TRUE, rep(FALSE, N - 1)), ], true))
      }
    }),
    PC = list(as.factor(1:10))
  ) %>%
  select(-c(fun, params, time, pcs)))

preci %>%
  unnest(cor, PC) %>%
  ggplot(aes(PC, cor^2, color = size, shape = size)) %>%
  bigstatsr:::MY_THEME(coeff = 2) +
  geom_point(cex = 5) +
  geom_line(aes(group = size, linetype = size), lwd = 2) +
  facet_wrap(~method, nrow = 2) +
  labs(x = "Principal Component", y = "Squared correlation between estimated and true PCs",
       color = "Number of individuals", shape = "Number of individuals", linetype = "Number of individuals") + 
  theme(legend.position = c(0.8, 0.15),
        strip.text.x = element_text(size = rel(3)),
        legend.key.width = unit(4.5, "line")) + 
  coord_cartesian(ylim = c(0, 1))

ggsave("figures/precision-pca.png", width = 1200, height = 900, scale = 1/75)
```

```{r, include=FALSE}
library(gmailr)
send_message(
  mime(To = "florian.prive.21@gmail.com",
       From = "florian.prive.21@gmail.com",
       Subject = sprintf("Knitting of %s has finished",
                         knitr::current_input()))
)
```