### 2018-01-18 GWAS-comparison.Rmd
save

---
title: "Genome-Wide Association Study"
subtitle: "Comparing bigstatsr and bigsnpr with PLINK and EIGENSOFT"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", 
                      fig.width = 16, fig.asp = 0.7)
options(width = 90)
```

## Prerequisites

Here, we suppose that the data have been pre-processed (quality control and imputation), and is available as a bk/rds (format of packages bigstatsr and bigsnpr). See the notebook "preprocessing".

```{r}
suppressMessages({
  # devtools::install_github("privefl/bigsnpr")
  library(bigsnpr)   
  # install.packages("glue")
  library(glue)      
  # install.packages("ggplot2")
  library(tidyverse)
})
# download.file("https://data.broadinstitute.org/alkesgroup/EIGENSOFT/EIG-6.1.4.tar.gz",
#               destfile = "EIG-6.1.4.tar.gz")
# untar("EIG-6.1.4.tar.gz")
smartpca <- "EIG-6.1.4/bin/smartpca"
plink <- download_plink()

write.table2 <- bigsnpr:::write.table2
fread2 <- function(...) data.table::fread(data.table = FALSE, ...)

NCORES <- nb_cores()
```

## Example of GWAS

We will show how to perform a GWAS with or without our packages. For the sake of reproducibility and simplicity, everything will be done inside R (via possible system calls to command-line software).

### With bigstastr and bigsnpr

```{r}
system.time(
  celiac <- snp_attach("backingfiles/celiacQC.rds")
)
G <- celiac$genotypes
CHR <- celiac$map$chromosome
POS <- celiac$map$physical.pos

# Pruning
system.time(
  ind.keep <- snp_pruning(G, infos.chr = CHR, 
                          exclude = snp_indLRLDR(CHR, POS),
                          ncores = NCORES)
)
```

```{r}
# PCA
system.time(
  obj.svd <- big_randomSVD(G, snp_scaleBinom(), 
                           ind.col = ind.keep,
                           ncores = NCORES)
)

# Get population from external files
pop.files <- list.files(path = "../thesis-celiac/Dubois2010_data/",
                        pattern = "cluster_*",
                        full.names = TRUE)
pop <- snp_getSampleInfos(celiac, pop.files)[[1]]
pop.names <- c("Netherlands", "Italy", "UK1", "UK2", "Finland")
# Plot PC1 & PC2
plot(obj.svd, type = "scores", coeff = 2) +
  aes(color = pop.names[pop]) +
  labs(color = "Population", title = NULL) +
  theme(legend.position = c(0.2, 0.75)) + 
  guides(colour = guide_legend(override.aes = list(size = 4)))

ggsave("figures/celiac-pca.png", width = 1124, height = 945, scale = 1/90)
```

```{r}
# GWAS
system.time(
  gwas <- big_univLogReg(G, y01.train = celiac$fam$affection - 1,
                         covar.train = obj.svd$u, ncores = NCORES)
)

labels <- 1:22; labels[c(15, 17, 19, 21)] <- ""
snp_manhattan(snp_gc(gwas), CHR, POS, 
              coeff = 2, labels = labels) + 
  geom_hline(yintercept = -log10(5e-8), col = "red") +
  labs(title = NULL) +
  coord_cartesian(ylim = c(0, 30))

ggsave("figures/celiac-gwas-cut.png", width = 1475, height = 788, scale = 1/90)
```

### With PLINK and EIGENSOFT (and R)

```{r}
# Pruning
prefix_bed <- tempfile()
snp_writeBed(celiac, bedfile = paste0(prefix_bed, ".bed"))
write.table2(LD.wiki34, "LRLDR.txt")
system.time(
  system(glue("{plink} --bfile {prefix_bed} --out {prefix_bed}",
              " --exclude LRLDR.txt --range",
              " --indep-pairwise 50 1 0.2",
              " --threads {NCORES}"))
)

ind.keep2 <- match(scan(paste0(prefix_bed, ".prune.in"), what = ""), celiac$map$marker.ID)
length(intersect(ind.keep, ind.keep2)) / length(union(ind.keep, ind.keep2))

# Make a pruned dataset
write.table2(as.data.frame(celiac$map[ind.keep, "marker.ID"]), "keep.in")
system.time(
  system(glue("{plink} --bfile {prefix_bed} --out {prefix_bed}",
              " --extract keep.in --make-bed",
              " --threads {NCORES}"))
)

# PCA with EIGENSOFT
fastpca <- function(prefix, fast = TRUE) {

  parfile <- glue("{prefix}.pca.par")

  ## Renaming so that Eigensoft can recognize file type
  system(glue("cp {prefix}.bim {prefix}.pedsnp"))  
  ## Renaming so that Eigensoft can recognize file type
  system(glue("cp {prefix}.fam {prefix}.pedind"))  
  writeLines(glue(
    "genotypename: {prefix}.bed
    snpname: {prefix}.pedsnp
    indivname: {prefix}.pedind
    evecoutname: {prefix}.pca.evec
    evaloutname: {prefix}.eval
    altnormstyle: NO
    numoutevec: 10
    numoutlieriter: 0
    numoutlierevec: 10
    outliersigmathresh: 6
    qtmode: 0
    fastmode: {as.integer(fast)}"
  ), con = parfile)
  ## PCA
  system(glue("{smartpca} -p {parfile}"), ignore.stdout = TRUE)
  ## removing temp files
  system(glue("rm {prefix}.pedsnp")) 
  system(glue("rm {prefix}.pedind"))
  system(glue("rm {parfile}"))

  evec <- data.table::fread(glue("{prefix}.pca.evec"), 
                            data.table = FALSE, skip = 1)

  as.matrix(evec[, 2:11])
}

system.time(
  obj.svd2 <- fastpca(tmp)
)

# FastPCA is not accurate enough (see Notebook "benchmark-pca")
round(100 * cor(obj.svd2, obj.svd$u), 1)
```

```{r}
# GWAS
write.table2(cbind(celiac$fam[, 1:2], obj.svd$u), "covar.txt")
system.time(
  system(glue("{plink} --bfile {prefix_bed} --out {prefix_bed}",
              " --logistic hide-covar",
              " --covar covar.txt",
              " --threads {NCORES}"))
)

gwas2 <- fread2(paste0(prefix_bed, ".assoc.logistic"))
all.equal(gwas2$P, predict(gwas, log10 = FALSE)) 
```