### 2018-01-18 proof-same-decomposition.Rmd

---
title: "Same eigen decomposition: proof"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Proof for the same pearson correlation coefficient

Let us note $x$ and $y$ two SNPs and $\overline{x}$ the mean of $x$.

Then the pearson correlation coefficient between $x$ and $y$ is: 
$$r = \dfrac{\sum\limits_{i = 1}^n (x_i - \overline{x}) (y_i - \overline{y})}{\sqrt{\sum\limits_{i = 1}^n (x_i - \overline{x})^2} \sqrt{\sum\limits_{i = 1}^n (y_i - \overline{y})^2}}~~.$$

If we note $x_K$ and $y_K$ the SNPs $x$ and $y$ replicated $K$ times, then, $\overline{x_K} = \frac{K \sum\limits_{i = 1}^n x_i}{K~n} = \frac{1}{n} \sum\limits_{i = 1}^n x_i = \overline{x}$ and $r_K$ the pearson correlation coefficient between $x_K$ and $y_K$ is 
$$r_K = \dfrac{K \sum\limits_{i = 1}^n (x_i - \overline{x_K}) (y_i - \overline{y_K})}{\sqrt{K \sum\limits_{i = 1}^n (x_i - \overline{x_K})^2} \sqrt{K \sum\limits_{i = 1}^n (y_i - \overline{y_K})^2}} = \dfrac{\sum\limits_{i = 1}^n (x_i - \overline{x}) (y_i - \overline{y})}{\sqrt{\sum\limits_{i = 1}^n (x_i - \overline{x})^2} \sqrt{\sum\limits_{i = 1}^n (y_i - \overline{y})^2}} = r~.$$

## Proof for the same eigen analysis

First, as $\overline{x_K}$ = $\overline{x}$, then the allele frequency of $x_K$ is the same as for $x$ so that the scaling $\frac{x - 2 p}{\sqrt{2 p (1 - p)}}$ is also the same (because $p = \overline{x} / 2$).

For the exact singular value decomposition of $G = U \Delta V^T$ (where $G$ is the scaled genotype matrix), we can first compute $\Sigma = G^T G = V \Delta^2 V^T$, then remark that $\Sigma V = V \Delta^2$ so that $V$ is the matrix of the eigen vectors of $\Sigma$ and $\Delta^2$ is the matrix of the eigen values of $\Sigma$. Finally to get $U$, we can compute $G V \Delta^{-1} = U \Delta V^T V \Delta^{-1} = U \Delta \Delta^{-1} = U$.

For replicated individuals, we want the decomposition of $G_K = U_K \Delta_K V_K^T$. Then, $\Sigma_K = G_K^T G_K = KG^T G = K \Sigma$ so that $V_K = V$ (same PC loadings) and $\Delta_K^2 = K \Delta^2$ resulting in $\Delta_K = \sqrt{K} \Delta$ (same eigen values, up to a constant). Finally $U_K = G_K V_K \Delta_K^{-1} = G_K V \left(\sqrt{K} \Delta\right)^{-1} = \frac{1}{\sqrt{K}} G_K V \Delta^{-1}$ (PCs scores are the same, up to a constant).