### 2018-01-23 GWAS-comparison.Rmd
GWAS + PRS

---
title: "Genome-Wide Association Study"
subtitle: "Comparing bigstatsr and bigsnpr with PLINK and EIGENSOFT"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", 
                      fig.width = 16, fig.asp = 0.7)
options(width = 110)
```

## Prerequisites

Here, we suppose that the data have been pre-processed (quality control and imputation), and is available as a bk/rds (format of packages bigstatsr and bigsnpr). See the notebook "preprocessing".

```{r}
suppressMessages({
  # devtools::install_github("privefl/bigsnpr")
  library(bigsnpr)   
  # install.packages("glue")
  library(glue)      
  # install.packages("ggplot2")
  library(tidyverse)
})
# download.file("https://data.broadinstitute.org/alkesgroup/EIGENSOFT/EIG-6.1.4.tar.gz",
#               destfile = "EIG-6.1.4.tar.gz")
# untar("EIG-6.1.4.tar.gz")
smartpca <- "EIG-6.1.4/bin/smartpca"
plink <- download_plink()

write.table2 <- bigsnpr:::write.table2
fread2 <- function(...) data.table::fread(data.table = FALSE, ...)

NCORES <- nb_cores()
```

## Example of GWAS

We will show how to perform a GWAS with or without our packages. For the sake of reproducibility and simplicity, everything will be done inside R (via possible system calls to command-line software).

### With bigstastr and bigsnpr

```{r}
system.time(
  celiac <- snp_attach("backingfiles/celiacQC.rds")
)
G <- celiac$genotypes
CHR <- celiac$map$chromosome
POS <- celiac$map$physical.pos

# Pruning
system.time(
  ind.keep <- snp_pruning(G, infos.chr = CHR, 
                          exclude = snp_indLRLDR(CHR, POS),
                          ncores = NCORES)
)
```

```{r}
# PCA
system.time(
  obj.svd <- big_randomSVD(G, snp_scaleBinom(), 
                           ind.col = ind.keep,
                           ncores = NCORES)
)

# Get population from external files
pop.files <- list.files(path = "../thesis-celiac/Dubois2010_data/",
                        pattern = "cluster_*",
                        full.names = TRUE)
pop <- snp_getSampleInfos(celiac, pop.files)[[1]]
pop.names <- c("Netherlands", "Italy", "UK1", "UK2", "Finland")
# Plot PC1 & PC2
plot(obj.svd, type = "scores", coeff = 2) +
  aes(color = pop.names[pop]) +
  labs(color = "Population", title = NULL) +
  theme(legend.position = c(0.2, 0.75)) + 
  guides(colour = guide_legend(override.aes = list(size = 4)))

ggsave("figures/celiac-pca.png", width = 1124, height = 945, scale = 1/90)
```

```{r}
# GWAS (logistic)
y01 <- celiac$fam$affection - 1
system.time(
  gwas <- big_univLogReg(G, y01.train = y01, covar.train = obj.svd$u, 
                         ncores = NCORES)
)
gwas.pval <- predict(gwas, log10 = FALSE)

labels <- 1:22; labels[c(15, 17, 19, 21)] <- ""
snp_manhattan(snp_gc(gwas), CHR, POS, coeff = 2, labels = labels) + 
  geom_hline(yintercept = -log10(5e-8), col = "red") +
  labs(title = NULL) +
  coord_cartesian(ylim = c(0, 30))

ggsave("figures/celiac-gwas-cut.png", width = 1475, height = 788, scale = 1/90)
```

```{r}
# GWAS (linear)
y <- rnorm(nrow(G))
system.time(
  gwas.lin <- big_univLinReg(G, y.train = y, covar.train = obj.svd$u, 
                             ncores = NCORES)
)
gwas.lin.pval <- predict(gwas.lin, log10 = FALSE) 
```

### With PLINK and EIGENSOFT (and R)

```{r}
# Write temporary PLINK files
prefix_bed <- tempfile()
bed <- snp_writeBed(snp_attach("backingfiles/celiacQC.rds"), 
                    bedfile = paste0(prefix_bed, ".bed"))
```

```{r}
# Long-range LD regions
write.table2(LD.wiki34, "LRLDR.txt")
# Pruning
system.time(
  system(glue("{plink} --bfile {prefix_bed} --out {prefix_bed}",
              " --exclude LRLDR.txt --range",
              " --indep-pairwise 50 1 0.2",
              " --threads {NCORES}"))
)

ind.keep2 <- match(scan(paste0(prefix_bed, ".prune.in"), what = ""), celiac$map$marker.ID)
# Same sets of SNPs
length(intersect(ind.keep, ind.keep2)) / length(union(ind.keep, ind.keep2))
```

```{r}
# Make a pruned dataset
write.table2(as.data.frame(celiac$map[ind.keep, "marker.ID"]), "keep.in")
system.time(
  system(glue("{plink} --bfile {prefix_bed} --out {prefix_bed}_sub",
              " --extract keep.in --make-bed",
              " --threads {NCORES}"))
)
```

```{r}
# PCA with EIGENSOFT
fastpca <- function(prefix, fast = TRUE) {

  parfile <- glue("{prefix}.pca.par")

  ## Renaming so that Eigensoft can recognize file type
  system(glue("cp {prefix}.bim {prefix}.pedsnp"))  
  ## Renaming so that Eigensoft can recognize file type
  system(glue("cp {prefix}.fam {prefix}.pedind"))  
  writeLines(glue(
    "genotypename: {prefix}.bed
    snpname: {prefix}.pedsnp
    indivname: {prefix}.pedind
    evecoutname: {prefix}.pca.evec
    evaloutname: {prefix}.eval
    altnormstyle: NO
    numoutevec: 10
    numoutlieriter: 0
    numoutlierevec: 10
    outliersigmathresh: 6
    qtmode: 0
    fastmode: {as.integer(fast)}"
  ), con = parfile)
  ## PCA
  system(glue("{smartpca} -p {parfile}"), ignore.stdout = TRUE)
  ## removing temp files
  system(glue("rm {prefix}.pedsnp")) 
  system(glue("rm {prefix}.pedind"))
  system(glue("rm {parfile}"))

  evec <- data.table::fread(glue("{prefix}.pca.evec"), 
                            data.table = FALSE, skip = 1)

  as.matrix(evec[, 2:11])
}

system.time(
  obj.svd2 <- fastpca(paste0(prefix_bed, "_sub"))
)

# FastPCA is not accurate enough (see notebook "benchmark-pca")
round(100 * cor(obj.svd2, obj.svd$u), 1)
```

```{r}
# GWAS
write.table2(cbind(celiac$fam[, 1:2], obj.svd$u), "covar.txt")
system.time(
  system(glue("{plink} --bfile {prefix_bed} --out {prefix_bed}",
              " --logistic hide-covar",
              " --covar covar.txt",
              " --threads {NCORES}"))
)

gwas2 <- fread2(paste0(prefix_bed, ".assoc.logistic"))
all.equal(gwas2$P, gwas.pval) 
```

```{r}
# GWAS (linear)
write.table2(cbind(celiac$fam[, 1:2], y), "qt.phe")
system.time(
  system(glue("{plink} --bfile {prefix_bed} --out {prefix_bed}",
              " --linear hide-covar --pheno qt.phe",
              " --covar covar.txt",
              " --threads {NCORES}"))
)

gwas2.lin <- fread2(paste0(prefix_bed, ".assoc.linear"))
all.equal(gwas2.lin$P, gwas.lin.pval) 
```

### With R packages SNPRelate and GWASTools

```{r}
library(SNPRelate)

# Transform files in our format and attach them
system.time(
  gds <- snpgdsBED2GDS(bed.fn = bed, 
                       fam.fn = sub("\\.bed$", ".fam", bed),
                       bim.fn = sub("\\.bed$", ".bim", bed),
                       out.gdsfn = paste0(prefix_bed, ".gds"),
                       cvt.snpid = "int")
)
```

```{r}
system.time(
  genofile <- snpgdsOpen(gds)
)
```

```{r}
# Pruning
system.time(
  snpset <- snpgdsLDpruning(genofile, method = "corr", ld.threshold = 0.2,
                            num.thread = NCORES)
) # 38 sec
ind.keep3 <- unlist(snpset, use.names = FALSE)
# Not the same set of SNPs (because not the same algo)
length(intersect(ind.keep3, ind.keep2)) / length(union(ind.keep3, ind.keep2))
```

```{r}
snpgdsClose(genofile)
# Need sudo yum install netcdf-devel.x86_64
library(GWASTools)
(genofile <- GdsGenotypeReader(gds, YchromCode=24L, XYchromCode=25L))
scanID <- getScanID(genofile)
# sex <- getVariable(genofile, "sample.annot/sex")
phenotype <- getVariable(genofile, "sample.annot/phenotype")
covar <- obj.svd$u
colnames(covar) <- paste0("PC", 1:10)
scanAnnot <- ScanAnnotationDataFrame(
  cbind.data.frame(scanID, pheno = phenotype - 1, covar)
)
snpID <- getSnpID(genofile)
chromosome <- getChromosome(genofile)
position <- getPosition(genofile)
alleleA <- getAlleleA(genofile)
alleleB <- getAlleleB(genofile)
snpAnnot <- SnpAnnotationDataFrame(data.frame(
  snpID = seq_along(snpID), chromosome, position, rsID = snpID,
  alleleA, alleleB, stringsAsFactors=FALSE),
  YchromCode=24L, XYchromCode=25L)

genoData <- GenotypeData(genofile, scanAnnot = scanAnnot, snpAnnot = snpAnnot)

# Logistic regression
system.time(
  gwas3 <- assocRegression(genoData, "pheno", 
                           model.type = "logistic",
                           covar = colnames(covar),
                           snpEnd = 10e3)
) * ncol(G) / 10e3
all.equal(gwas3$Wald.pval, gwas.pval[rows_along(gwas3)]) 
# assocRegression uses glm() so that it is the most precise
# We are very precise and PLINK is the least precise
true.pval <- sapply(1:9, function(j) {
  summary(glm(y01 ~ G[, j] + covar, family = "binomial"))$coefficients[2, 4]
})
rbind(`true (glm)` = true.pval, bigstatsr = gwas.pval[1:9], 
      plink = gwas2$P[1:9], gwastools = gwas3$Wald.pval[1:9])
```

```{r}
scanAnnot <- ScanAnnotationDataFrame(
  cbind.data.frame(scanID, pheno = y, covar)
)

genoData <- GenotypeData(genofile, scanAnnot = scanAnnot, snpAnnot = snpAnnot)

# Linear regression
system.time(
  gwas3.lin <- assocRegression(genoData, "pheno", 
                               model.type = "linear",
                               covar = colnames(covar),
                               snpEnd = 10e3)
) * ncol(G) / 10e3
all.equal(gwas3.lin$Wald.pval, gwas.lin.pval[rows_along(gwas3)])

# Again, we are very precise
true.lin.pval <- sapply(1:9, function(j) {
  summary(lm(y ~ G[, j] + covar))$coefficients[2, 4]
})
rbind(`true (lm)` = true.lin.pval, bigstatsr = gwas.lin.pval[1:9], 
      plink = gwas2.lin$P[1:9], gwastools = gwas3.lin$Wald.pval[1:9])
```

```{r}
# Cleaning
file.remove(c("LRLDR.txt", "covar.txt", "keep.in", "qt.phe"))
```