### 2018-01-23 PRS-comparison.Rmd

---
title: "Polygenic Risk Scores (PRS)"
subtitle: "Comparing bigstatsr and bigsnpr with PRSice"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", 
                      fig.width = 8, fig.asp = 0.7)
options(width = 110)
```

## Prerequisites

Here, we suppose that the data have been pre-processed (quality control and imputation), and is available as a bk/rds (format of packages bigstatsr and bigsnpr). See the notebook "preprocessing".

```{r}
suppressMessages({
  # devtools::install_github("privefl/bigsnpr")
  library(bigsnpr)   
  # install.packages("glue")
  library(glue)      
  # install.packages("ggplot2")
  library(tidyverse)
})
# install.packages("data.table")
fread2 <- function(...) data.table::fread(data.table = FALSE, ...)
NCORES <- nb_cores()
```

We also need some summary statistics. Let's use ones for the height.

```{r, eval=FALSE}
download.file("http://portals.broadinstitute.org/collaboration/giant/images/0/01/GIANT_HEIGHT_Wood_et_al_2014_publicrelease_HapMapCeuFreq.txt.gz",
              destfile = "sumstats-height.txt.gz")
R.utils::gunzip("sumstats-height.txt.gz")
```

## Example of Polygenic Risk Scores (PRS)

We will show how to perform a GWAS with or without our packages. For the sake of reproducibility and simplicity, everything will be done inside R (via possible system calls to command-line software).

### With PRSice v2.1.0.beta

```{r}
# Write temporary PLINK files
prefix_bed <- tempfile()
bed <- snp_writeBed(snp_attach("backingfiles/celiacQC.rds"), 
                    bedfile = paste0(prefix_bed, ".bed"))
```

```{r}
# download.file("https://github.com/choishingwan/PRSice/releases/download/2.1.0.beta/PRSice_linux.zip",
#               destfile = "PRSice_linux.zip")
# unzip("PRSice_linux.zip", files = c("PRSice.R", "PRSice_linux"))
system("chmod +x PRSice_linux")
# Use PRSice
system.time(
  system(glue("Rscript PRSice.R",
              " --prsice PRSice_linux",
              " --base sumstats-height.txt",
              " --target {prefix_bed}",
              " --out {prefix_bed}",
              " --thread {NCORES}",
              " --pvalue p",
              " --stat b",
              " --beta T",
              " --snp MarkerName",
              " --A1 Allele1 --A2 Allele2",
              " --binary-target T"))
)
# Read results
file.prsice <- paste0(prefix_bed, ".prsice")
res.prsice <- fread2(file.prsice)
```

### With bigstastr and bigsnpr

```{r}
system.time(
  celiac <- snp_attach("backingfiles/celiacQC.rds")
)
G <- celiac$genotypes
CHR <- celiac$map$chromosome
POS <- celiac$map$physical.pos
```

```{r}
# Read the summary stats
system.time(
  sumstats <- fread2("sumstats-height.txt")
)
```

```{r}
system.time({
  # Get the SNP indices that correspond in sumstats
  indSNP <- match(celiac$map$marker.ID, sumstats$MarkerName)
  betas <- sumstats$b[indSNP]
  pval <- sumstats$p[indSNP]
  lpS <- -log10(pval)
  # And check if the reference alleles are switched
  same <- same_ref(celiac$map$allele1, celiac$map$allele2, 
                   sumstats$Allele1[indSNP], sumstats$Allele2[indSNP])
})
# Non-matching SNPs
sum(is.na(indSNP))
```

```{r}
# Use the same p-value thresholds
thrs <- -log10(res.prsice$Threshold)
# Clumping
system.time(
  ind.keep <- snp_clumping(G, infos.chr = CHR, S = lpS, thr.r2 = 0.1,
                           size = 250, is.size.in.bp = TRUE, infos.pos = POS,
                           exclude = which(is.na(same) | (lpS < min(thrs))), 
                           ncores = NCORES)
)
```
```{r}
# PRS
system.time(
  prs <- snp_PRS(G, betas.keep = betas[ind.keep],
                 ind.test = rows_along(G),
                 same.keep = same[ind.keep], 
                 lpS.keep = lpS[ind.keep], 
                 ind.keep = ind.keep,
                 thr.list = thrs)
)
```

```{r}
# p-values
system.time(
  pval.bigsnpr <- prs %>%
    big_copy() %>%
    big_univLogReg(celiac$fam$affection - 1) %>%
    predict(log10 = FALSE)
)
```

### Verification

```{r}
# p-values
cor(log10(pval.bigsnpr), log10(res.prsice$P))
bigstatsr:::MY_THEME(qplot(pval.bigsnpr, res.prsice$P, log = "xy")) +
  geom_abline(slope = 1, intercept = 0, col = "red") +
  labs(x = "P-values returned with bigsnpr (log-scale)",
       y = "P-values returned with PRSice (log-scale)")
```

```{r}
# Number of SNPs for each threshold
nb.pred <- sapply(thrs, function(thr) sum(lpS[ind.keep] > thr, na.rm = TRUE))
bigstatsr:::MY_THEME(qplot(nb.pred, res.prsice$Num_SNP)) +
  geom_abline(slope = 1, intercept = 0, col = "red") +
  labs(x = "Number of SNPs with bigsnpr",
       y = "Number of SNPs with PRSice")
```

These small differences are due to the way LD is assessed in the clumping algorithm.